\textit{Github :}

Github offers one of the most comprehensive forge APIs with many endpoints and features. Two APIs are proposed, a REST and a GraphQL API, enabling an easy access to information on a given element of the model used by Github (e.g., information on a repository, on a project's or a user's output, etc.). Search endpoints are available and offer Repository, Commit or Code searches on all Github.com. 
At first sight we might assume that the APIs of Github would allow us to implement our running request. 
We propose to split the query in two subquery that can be converted to GitHub API requests :
\begin{enumerate}
\item \textit{Subquery 1 (SQ1) "repositories of android applications"} can be fulfilled using the code search REST API to retrieve all the file having AndroidManifest.xml as name and their corresponding repository.
\begin{lstlisting}
https://api.github.com/search/code?q=Q 
//With the following query
Q=filename:AndroidManifest.xml
\end{lstlisting}


\item \textit{Subquery 2 (SQ2) "repositories which have a creation date not prior to 2015 and at least 1000 commits"} can be handled by leveraging on the GraphQL API to get the creation date and the total number of commit of each repository return in 1. (NB: the REST API might also be used but will need two requests instead of one for the same result)
\begin{lstlisting}
query { 
 repository(owner: "OWNER", name: "REPOSITORY") {
          createdAt
          defaultBranchRef {
            target {
              ... on Commit {
                history {
                  totalCount
                }
              }
            }
        }
	}
}
\end{lstlisting}
\end{enumerate}

Thus we show that the Github API'S are expressive enough to allow us to handle our query. However, there are many limitations that make this task difficult to implement.

\underline{Rate Limitation} : The second limitation is the number of API call and its impact in term of API rate limit. Indeed, GitHub has a fixed rate limit for each user and we did not find any paid plan to increase this rate limit. 
\begin{itemize}
\item The first subquery SQ1, between 3 and 9 million results are expected. Knowing that at most 100 results are returned per request, it would take between 30K and 90K queries. The rate limit is 30 queries per minute for the search endpoint, consequently it would take between 17h and 50h to execute SQ1.
\item The second subquery SQ2, is then called on all the distinct Repository return  by SQ1. As multiple AndroidManifest.xml can be contained in the same repository the number of repository returned will be less important than the number of file return by the api call. Nevertheless million of repository will be return. The second subquery SQ2, levarage on the graphql api that have a rate limit of 5000 requests per hour which means that 4 million graphql API call will take more than 32 days. 
\end{itemize} 
All the previous query execution time estimations are based on the assumption that rate limits are applicable in practice. However, in  we have not been able to use the entire quota. During our experiment we faced, like many users\footnote{\url{https://github.com/orgs/community/discussions?discussions_q=secondary+rate+limit}}, "secondary rate limitation" that are normally raised when an unappropriated use of the API is performed. However, there is no apparent reason given that we respect the Github best practices (e.g. waiting at least 1s between queries, respect rate limitation, use appropriate headers ...) following \footnote{\url{https://docs.github.com/en/rest/guides/best-practices-for-integrators?apiVersion=2022-11-28}}. According to the documentation, in this case, it is necessary to retry after the new rate limit reset even if not all the quota has been consumed, drastically reducing the number of allowed requests (1 request per minute instead of 30 in the worst case). These problems make it almost impossible to implement SQ2, which would take months to execute.
Finally, we observed that the \textit{\url{per_page}} attribute setting the number of entries per page returned and therefore per request does not always behave as expected, as the API return in some cases less elements than expected, requiring more requests to find all the entries.
\underline{Operationalization limitation} : The second limitation is related to the difficulty of operating the CODE search endpoint on such massive queries. The search endpoint only handles query that have less than a thousand of element in response \footnote{https://docs.github.com/en/rest/search?apiVersion=2022-11-28}. Our query return millions of result and is therefore not usable. A common tweak is to sample massive query using available attribute filter (e.g a query that target the Repository search endpoint can be sampled using repository creation date), such treatment need complex heuristic which makes its implementation very constraining. The only attribute that can be used to sample SQ1 is the file size. In addition to the complexity of the sampling heuristic, we observe that the Search endpoint is not deterministic and not reliable. We reduce our query to AndroidManifest.xml having a size in [100000,110000] bytes in order to retrieve less than a thousand of results. 
\begin{lstlisting}
curl -i \
  -H "Accept: application/vnd.github+json" \
  -H "Authorization: Bearer ghp_..."\
    'https://api.github.com/search/code?q=filename:AndroidManifest.xml+size:100000..110000&per_page=1'
\end{lstlisting}
It appears that the total count of results differ between call of the same query. Leading to non-reproducible results. 
\begin{lstlisting}
//Attempt#1 :
"total_count":283
//Attempt#2 :
"total_count": 317,
//Attempt#3 :
"total_count": 327,
\end{lstlisting}
All these limitations make the use of this API difficult for large scale mining purposes, both in terms of usability and the time needed to perform the requests.
\\
\\
\textit{Gitlab} 
\\
\hl{Gitlab} is an open source forge and an end-to-end software development platform allowing code version control, continuous integration or issue tracking. The Gitlab API allow users to manipulate gitlab ressources (projects,users ..) and perform CRUD operations. The Gitlab API can also be used to automate Gitlab from external resources or within GitlabCI. Similarly to GitHub, Gitlab offers a legacy REST Api as well as a graphQl API. Search endpoints are available, allowing search for various type of items such as commit or project in a Gitlab instance. 

\underline{Search Endpoints limitation}
The Gitlab advanced search API \footnote{\url{https://docs.gitlab.com/ee/api/search.html}} allows premium user to search for files by name through the "blob" scope across the entire gitlab.com instance, according to the documentation.However, this feature is not available on gitlab.com, so that the first SBQ1 subquery is not supported by the Gitlab API\footnote{\url{https://gitlab.com/gitlab-org/gitlab/-/issues/197231}}.

Although it is not possible to search directly through all projects in the instance, it is possible to check for a given project if it contains an "AndroidManifest.xml" file.  :
\begin{itemize}
    \item The "GET /projects/:id/repository/tree" endpoint of the api can be used to list the contents of a repository tree. This requires a lot of pagination related queries. A trick is to limit the number of queries by interjecting an undocumented internal api, used in the gitlab web interface, GET https://gitlab.com/gitlab-org/gitlab/-/files/master?format=json.
    \item The legacy search api can be used, as it only allows searching for files from their content, it is necessary to identify a sub string of the content to isolate this file ( eg "<manifest")
\end{itemize}





Also, targeting a global scope for code search on GitLab.com won’t work at the moment 2

search api 10/minutes
else 300/minutes ->432k par jour

"Search query, which can be for the project name, a path, or a description."



Projects api to iterate over project
https://docs.gitlab.com/ee/api/projects.html
https://docs.gitlab.com/ee/api/index.html#keyset-based-pagination
not possible to know the results size if a query return more than 10k records 

No clear information about how much repo are in gitlab.com however SWH crawl 3.9 M of repositories. https://archive.softwareheritage.org/
Gitlab has a search endpoint but it do not allow to search file on the entire gitlab instance \footnote{https://forum.gitlab.com/t/advanced-search-not-searching-code-using-gold-trial/44149
}. However, we can do a search on a single repository.

We cannot get first all the repositories containing a file named "AndroidManifest.xml" and then filter those having a creation date not prior to 2015 and at least 1000 commits. We need to do the process in the opposite order.

The gitlab api allow retrieving 100 records per page, more than 39000 API call are needed to get all the repository information. The rate limit for this request is 2000 requests per minute, 19.5 minutes are therefore needed. This query return information about repository, including the creation date, allowing us to filter those created after 2015. A second call is needed to get the total commit number per repositories. 

heure 

Gitlab search api 





A graph query langage is needed to express our query on the SWH graph property dataset. 

Our objective being to compile the requests towards the java swh-graph api, we also took into account the language tooling enabling to achieve a generator (e.g Does the metamodel is available or just the grammar ? Does it support model linking ). 
\begin{itemize}
    \item GraphQL : GraphQL is a query language that allows user to query data and choose the structure of the response. GraphQL is not designed to specify graph traversal query, for instance the transitive closure is not a concept of the language. There are some expressivity limitation such as the lack of support of conditional, user defined function, intermediate variable or join operation. The processing must therefore be done on the client's side, which can be an issue in a context of massive graph and big data. 
    \item Cypher : Cypher is an open query language for property graph databases, allowing to manipulate and query graph elements in an intuitive way. Cypher addresses most of the limitations mentioned for graphql except user defined function making it a good candidate as our query description language. 
\end{itemize}
cypher limitation 
- no built-in support for user defined function
- limited expressiveness
